---
title: "HW 4"
author: "Denys Osmak"
date: "2024-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
# importing libs 
library(rvest)
library(sas7bdat)
library(tibble)
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
library(lubridate)
library(knitr)
library(mosaic)
library(kableExtra)
```

# Problem 1

```{r}
# importing the data
pr1Data = read.csv("gasprices.csv")
```

### Gas stations charge more if they lack direct competition in sight.

#### Evidence:

```{r}
# bootsrtap the diffmean 10k times  
boot_1_1 = do(10000) * diffmean(Price ~ Competitors , data=resample(pr1Data) )


```

```{r}
# just some data analysis for better understanding of the dataset
mean(Price ~ Competitors , data=pr1Data)
                      
ggplot(boot_1_1, aes(x=diffmean) ) + geom_histogram() + labs(title = "Difference of Price Gas stations with & without competitors")

# grabing the confidence level 
confint(boot_1_1, method = "percentile")

```

#### Conclusion:

We can tell that this claim is true and supported by the data. In our data set, on average the gas station that did not have any competition nearby had charged 0.02 dollars more per gas vs gas stations with nearby competitors. Applying bootstrapping, with 95% accuracy we can articulate that the wider population of gas stations are charging anywhere from 0.05 more or to 0.01 less per gallon.

### The richer the area, the higher the gas prices

#### Evidence:

```{r}
# bootsrtap the diffmean 10k times 
boot_1_2 = do(10000) * mean(Price ~ Income , data=resample(pr1Data) )

```

```{r}
# just some data analysis for better understanding of the dataset
mean(Price ~ Income , data=pr1Data)

sum1_2 = group_by(pr1Data, Income) %>%
  summarise(mean(Price))

sum1_22 = group_by(boot_1_2, Income) %>%
  summarise(mean(Price))
  
                      
ggplot(pr1Data, aes(x=Income, y=Price) ) + geom_point() + labs(title = "ORIGINAL DATA Income vs Gas Price")

ggplot(sum1_2, aes(x=Income, y="mean(Price)") ) + geom_point() + labs(title = "SUM DATA Income vs Gas Price") 

ggplot(boot_1_2, aes(x=Income, y=Price) ) + geom_point() + labs(title = "Income vs Gas Price")


# grabing the confidence level 
confint(boot_1_1, method = "percentile")


```

#### Conclusion:

### Gas stations at stoplights charge more.

#### Evidence:

```{r}
# bootsrtap the diffmean 10k times  
boot_1_3 = do(10000) * diffmean(Price ~ Stoplight , data=resample(pr1Data) )

```

```{r}
# just some data analysis for better understanding of the dataset
mean(Price ~ Stoplight , data=pr1Data)
                      
ggplot(boot_1_3, aes(x=diffmean) ) + geom_histogram() + labs(title = "Difference of Price Gas stations with & without Stoplights")

# grabing the confidence level 
confint(boot_1_3, method = "percentile")


```

#### Conclusion:

While the data collected on the 101 gas stations shows an average increase in price by 0.6 cents the data does not prove that the gas stations next to stoplights charge more. We can claim with 95% confidence that gas stations next to stoplights can either charge 4 cents lower or 3 cents higher then gas stations away from stoplights. Meaning the results are not statistically significant.

### Gas stations with direct highway access charge more.

#### Evidence:

```{r}
# bootsrtap the diffmean 10k times  
boot_1_4 = do(10000) * diffmean(Price ~ Highway , data=resample(pr1Data) )

```

```{r}
# just some data analysis for better understanding of the dataset
table( pr1Data$Highway)
mean(Price ~ Highway , data=pr1Data)
diffmean(Price ~ Highway , data=pr1Data)
                      
ggplot(boot_1_4, aes(x=diffmean) ) + geom_histogram() + labs(title = "Difference of Price Gas stations with & without Highway access")

# grabing the confidence level 
confint(boot_1_4, method = "percentile")


```

#### Conclusion:

From the collected 101 gas stations, the gas stations with highway access charged 4.5 cents more then gas stations without highway access. However, we can say with 95% confidence that gas stations with highway access can charge anywhere from 0.9 to 8 cents more then gas stations without highway access, making this a statistically significant find, and a practically significant find.

### Shell charges more than all other non-Shell brands.

#### Evidence:

```{r}
# mutate the data to have Y and N if it is a shell compnay

pr1Data = mutate(pr1Data, ShellBool = pr1Data$Name == "Shell")

# bootstrap 10k times 

boot_1_5 = do(10000) * diffmean(Price ~ ShellBool , data=resample(pr1Data) )

```

```{r}
# just some data analysis for better understanding of the dataset
mean(Price ~ ShellBool , data=pr1Data)
diffmean(Price ~ ShellBool , data=pr1Data)
table( pr1Data$ShellBool)
                      
ggplot(boot_1_5, aes(x=diffmean) ) + geom_histogram() + labs(title = "Difference of Price Shell Gas Stations VS non Shell Gas Stations")

# grabing the confidence level 
confint(boot_1_5, method = "percentile")

```

#### Conclusion:

From the collected 101 gas stations, the Shell brand gas stations with highway access charged 2.7 cents more then other brand gas stations However, we can say with 95% confidence that Shell gas stations can charge anywhere from 1 cent less to 6.4 cents more then other brand gas stations, making this a statistically insignificant find, yet a practically significant find.

# Problem 2

```{r}
pr2Data = read.csv("sclass.csv")

# get some insight on the data 

table(pr2Data$condition)
```

## Part A

```{r}

# filter by year 2011 & trim == "63 AMG"

pr2ptA = filter(pr2Data, pr2Data$year == 2011 & pr2Data$trim == "63 AMG")

# Get some data based on avaraged millage 

ggplot(pr2ptA, aes(x=mileage) ) + geom_histogram() 

# bootsrap 10k based on these cars 
pr2PtABoot = do(10000) * mean(x=sample(pr2ptA$mileage,116,replace = TRUE), data=replace(pr2ptA))

ggplot(pr2PtABoot, aes(x=mean) ) + geom_histogram() + labs(title = "Bootsraped Distribution of AVG Millage", x="Avarage Millage")

# get confidance interval
confint(pr2PtABoot, method = "percentile")



```

### Conclusion:

**The average millage of 2011 63 AMG that hit the market is 28,997 miles +- 2,712.19 miles**

## Part B

```{r}
# filter by year 2014 & trim == "550" %>% mutate isBlack 

pr2ptB = filter(pr2Data, pr2Data$year == 2014 & pr2Data$trim == "550") %>% 
  mutate(., isBlack = .$color == "Black")

# Get some data based on proportion on the entire subset 

pr2PtBTable = table(pr2ptB$isBlack)
pr2PtBTable
pr2PtBProp = count(pr2ptB$isBlack) / 2889

# bootsrap 10k based on these cars 
pr2PtBBoot = do(10000) * (count(sample(pr2ptB$isBlack,2889,replace = TRUE)) / 2889)




```

```{r}

ggplot(pr2PtBBoot, aes(x=n_TRUE) ) + geom_histogram() + labs(title = "Bootsraped Distribution of Black Car Proportions", x="Proportion of Black Cars")

# get confidance interval
confint(pr2PtBBoot, method = "percentile")
```

### Conclusion

**I can say with 95% confidence interval that on average 43.475% +-1.834% of all 2014 Mercedes S Class 550 on the used market are Black.**

# Problem 3

```{r}
pr3Data = read.csv("nbc_pilotsurvey.csv")

```

## Part A: Does "Living with Ed" or "My name is Earl" make people happier

### Approach: 

### Evidence:

```{r}

# filter all data to living with ed and my name is earl only

pt3A = filter(pr3Data, pr3Data$Show == "Living with Ed" | pr3Data$Show == "My Name is Earl" )

# find the mean for the overall data 

pt3A %>% 
  group_by(Show)%>% 
  table(pt3A$Q1_Happy)

# bootsrtrap 10k times 

```

```{r}

# show diffmean hist 


# show box plot of the shows 


# confidence interval


```

### Conclusion:

## Part B: 

### Approach:

### Evidence:

### Conclusion:

## Part C: 

### Approach:

### Evidence:

### Conclusion:

# Problem 4

```{r}
pr4Data = read.csv("ebay.csv")

```
