---
title: "HW 2"
author: "Denys Osmak"
date: "2024-01-25"
output:
  pdf_document: default
knitr:
  opts_chunk:
    warning: no
    echo: no
    message: no
---

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(rvest)
library(sas7bdat)
library(tibble)
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
library(lubridate)
library(knitr)
```

# Github link:

<https://github.com/DenysUkr/SDS-315-GitHub-Repo/tree/main/Homework/HW%202>

# Problem 1

```{r warning=FALSE, echo=FALSE, message=FALSE}
# importing the data
pr1Data = read.csv("profs.csv")

```

## Part A

```{r warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data=pr1Data) + geom_histogram(aes(eval),binwidth = .1 , fill="red4", col="black") + labs(title="Distribution of Course Evaluation Scores", x = "Evalation Score")

```

We can tell from this graph that the largest majarity of the professors are evaluated at around the 4/5 mark. While no professors are marked to be 2/5 and bellow.

## Part B

```{r warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data=pr1Data) + geom_boxplot(aes(x=native, y=eval) , fill="red4", col="black") + labs(title="Distribution of Course Evaluation Scores Depending on Native English Speaking Professors", y = "Evaluation Score", x="Native English Speaking Profesor") 

```

This box plot shows that there is a significant difference in evaluation scores between professors that are native English speakers vs that are not. From the graph we can see that on avarage a native English speaking professor gets \~ 0.5 higher score then a non native English speaker.

## Part C

```{r warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data=pr1Data) + geom_histogram(aes(x=eval, y=..density..),binwidth = .1 , fill="red4", col="black") + labs(title="Desity Histograms of Course Evaluation Scores vs Gender", x = "Evalation Score") + facet_wrap(vars(gender))

```

Due to the fact that there seems to be a larger number of male professors, to see the true difference in the distribution of the evaluation score between male and female professors, the charts were normalized. After normalizing the charts it can be seen that while the male and female professors have relativly same evaluation scores, most of them being between 3.5-4.5, there is an abnormal ammount of female professors that are rated 3.5 and male professors at 4.5.

## Part D

```{r warning=FALSE, echo=FALSE, message=FALSE}

ggplot(data=pr1Data) + geom_point(aes(x=beauty, y=eval), col="red4") + labs(title="Beauty Score vs Evaluation Score", x= "Beauty Score", y = "Evalation Score") 

```

From the scatter plot there seems to be little to no correlation between beauty score and the evaluation score.

# Problem 2

## Part A

```{r warning=FALSE, echo=FALSE, message=FALSE}
# importing data
pr2Data = read.csv("bikeshare.csv")

# finding the avarage boardings per hour of the day for hour
# Nov
pr2Data = pr2Data %>% 
  group_by(hr) %>%
  mutate(avgRentals = mean(total))


# creating lien graph

ggplot() + geom_line(data=pr2Data, aes(y=avgRentals, x=hr), col="red4") + labs(title="Hourly Bike Rentals vs Time", x = "Hour of Day", y="Avarage Bike rental") 


```

By looking at this graph we can conclude that there is a large spike of bike rantals during the rush hours around 7 am and 6pm when people are commuting to and from work.

## Part B

```{r warning=FALSE, echo=FALSE, message=FALSE}

pr2DataB = pr2Data %>% 
  group_by(hr,workingday) %>%
  mutate(avgRentals = mean(total))


ggplot(pr2DataB) + geom_line(aes(y=avgRentals, x=hr), col="red4") + labs(title="Hourly Bike Rentals vs Time", x = "Hour of Day", y="Avarage Bike rental") + facet_wrap(~workingday)
```

From the graph we can see that the weekends (identefied as 0) have a much more relaxed curve. This is due to the fact that during the weeked the bikes are used for pleasure rather then for commuting to work.

## Part C

```{r warning=FALSE, echo=FALSE, message=FALSE}

pr2DataC = pr2Data %>% 
  filter(hr == 9) %>% 
  mutate(avgRentals = mean(total))

ggplot(pr2DataC) + geom_col(aes(y=avgRentals, x=weathersit), col="red4", fill="black") + labs(title="Hourly Bike Rentals at 9am vs Weather Code", x = "Weather Code", y="Avarage Bike rental")

```

We can see that the average ridership goes down when the weather is cloudy or has light rain ( type 2 & 3 weather codes). Additionally there is no recorded ridership at 9am during heavy rains.

# Problem 3

## Part A

```{r warning=FALSE, echo=FALSE, message=FALSE}

#importing data
pr3Data = read.csv("capmetro_UT.csv")

pr3Data = mutate(pr3Data,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))


# finding the avarage boardings per hour of the day for each month 
# Nov
pr3DataN = pr3Data %>% 
  group_by(day_of_week, hour_of_day) %>%
  filter(month == "Nov") %>% 
  mutate(avgBoarding = mean(boarding))
# Oct 
pr3DataO = pr3Data %>% 
  group_by(day_of_week, hour_of_day) %>%
  filter(month == "Oct") %>% 
  mutate(avgBoarding = mean(boarding))
# Sep
pr3DataS = pr3Data %>% 
  group_by(day_of_week, hour_of_day) %>%
  filter(month == "Sep") %>% 
  mutate(avgBoarding = mean(boarding))

  
# creating a line graph 
ggplot() + geom_line(data=pr3DataN, aes(x=hour_of_day, y=avgBoarding, col="November")) + geom_line(data=pr3DataO, aes(x=hour_of_day, y=avgBoarding, col="October")) + geom_line(data=pr3DataS, aes(x=hour_of_day, y=avgBoarding, col="September")) + labs(title="Avarage Hourly Boarding per hour over 3 months", x = "Hour of Day", y="Avarage Boarings") + facet_wrap(vars(day_of_week)) + scale_color_manual(values = c("November" = "red4", "October" = "navy", "September" = "darkgreen"))


```

```{r}
########################### TESTING NEW MEATHODS ########################### 

# Tesiting out an all in one solution
pr3DataTesting = pr3Data %>% 
  group_by(day_of_week, hour_of_day, month) %>%
  summarise(avgBoarding = mean(boarding))


# creating a line graph 
ggplot(pr3DataTesting) + geom_line( aes(x=hour_of_day, y=avgBoarding, col="November"))  + labs(title="Avarage Hourly Boarding per hour over 3 months", x = "Hour of Day", y="Avarage Boarings") + facet_wrap(vars(day_of_week)) + scale_color_manual(values = c("November" = "red4", "October" = "navy", "September" = "darkgreen"))


ggplot() + geom_line( data=pr3DataN,aes(x=hour_of_day, y=avgBoarding, col="November"))  + labs(title="Avarage Hourly Boarding per hour over 3 months", x = "Hour of Day", y="Avarage Boarings") + facet_wrap(vars(day_of_week)) + scale_color_manual(values = c("November" = "red4", "October" = "navy", "September" = "darkgreen"))






ggplot() + geom_line(data=filter(pr3DataTesting, month == "Nov"), aes(x=hour_of_day, y=avgBoarding, col="November")) + geom_line(data=filter(pr3DataTesting, month == "Oct"), aes(x=hour_of_day, y=avgBoarding, col="October")) + geom_line(data=filter(pr3DataTesting, month == "Sep"),  aes(x=hour_of_day, y=avgBoarding, col="September")) + labs(title="Avarage Hourly Boarding per hour over 3 months", x = "Hour of Day", y="Avarage Boarings") + facet_wrap(vars(day_of_week)) + scale_color_manual(values = c("November" = "red4", "October" = "navy", "September" = "darkgreen"))

```

By looking at the graphs, we can tell that there is the peak for boarding for weekdays are around 3-6pm. The weekdays have very similar curves, no matter the day of the week possibly attributed because the buses are primary used by work commuters. Additional, the reason that November might have the lowest is because it is the coldest month. However the reason why September might have the lowest boarding on Mondays might be due to the fact that there was some sort of holiday on Monday of September of 2018, which lowered the avarage hourly boarding.

## Part B

```{r warning=FALSE, echo=FALSE, message=FALSE}

ggplot(data=pr3Data) + geom_point( aes(x=boarding, y = temperature, color=weekend)) + labs(title="Avarage Hourly Boarding per hour over 3 months", y = "Temperature (F)", x ="Total Boarings") + facet_wrap(vars(hour_of_day)) 


```

There is little to no correlation between the temperature and the transit usage, since there is a completely vertical slope as shown on the graph. Interesting enough the weekends tend to have less ovaral boarding.

# Problem 4

## Part A

```{r warning=FALSE, echo=FALSE, message=FALSE}

pr4Data = read.csv("billboard.csv")

# counting the numbers of weeks that all of the song were on bilbord top 100

pr4DataNew = pr4Data %>% 
  group_by(performer, song) %>%
  summarise(totalWeeks = max(weeks_on_chart)) %>% 
  arrange(desc(totalWeeks))


pr4DataTrimed = head(pr4DataNew, 10)
kable(pr4DataTrimed, caption = "Top Songs")
```

## Part B

```{r warning=FALSE, echo=FALSE, message=FALSE}
# filtering and remoing 58 & 21
pr4DataPartB = filter(pr4Data, year != 1958 & year != 2021)


pr4DataPartB = pr4DataPartB %>% 
  group_by(year) %>% 
  arrange(year) %>% 
  summarise(uniqueness = length(unique(song_id)))


ggplot(data=pr4DataPartB) + geom_point(aes(x=year, y=uniqueness), col="red4") + labs(title="How many unique songs there were each year in top 100", x= "Year", y = "Num of Songs") 
```

There is an interesting dip how number of unique songs have doped significantly from 1960 to 2000 and then rapidly picked back up after 2000. This might be to the fact of a music monopoly or some other extreme factor

## Part C

Let's define a "ten-week hit" as a single song that appeared on the Billboard Top 100 for at least ten weeks. There are 19 artists in U.S. musical history since 1958 who have had at least 30 songs that were "ten-week hits." Make a bar plot for these 19 artists, showing how many ten-week hits each one had in their musical career. Give the plot an informative caption in which you explain what is shown

## 
